---
title: "STAT 656: Project "
subtitle: ""
author: "Andrew Fillmore"
date: "3/19/2021"
output: html_document
---

```{r setup, include=FALSE,message=FALSE, warning=FALSE}
require(dplyr)
require(caret)
require(quantmod)
require(tidyverse)
require(tidyquant)
require(dwt)

rm(list=ls())
```


First we must get our data. We will look at the logarithmic returns for stocks in the S&P400 over the two year period ranging from January 1, 2019 to December 31, 2020.

```{r import_data, cache=TRUE}
returns <- tq_index('SP600') %>%
  tq_get(from='2019-01-01', to='2020-12-31', verbose=T) %>%
  group_by(symbol) %>%
  tq_transmute(select=adjusted, mutate_fun=periodReturn, period='daily', type='log', col_rename='log_daily_returns') %>%
  pivot_wider(names_from=date, values_from=log_daily_returns) %>%
  ungroup %>%
  drop_na
```
#Principle Components

Now we look at the Principle Components. This should give us a look at whether or not there are obvious groupings in the data.

```{r PCA_returns}
returns_PC <- prcomp(returns %>% select(where(is.numeric)), scale=F, center=T)
screeplot(returns_PC)
```

Much of the variance is explained in the first principle component, but the drop off does not suggest that we should be able to use the principle components inplace of the data for clustering.

```{r PCA_returns_plot}
returns_PC_df <- data.frame(returns_PC$x)
ggplot(data=returns_PC_df, aes(x=PC1, y=PC2)) + 
  geom_point() +
  coord_cartesian(xlim=range(returns_PC_df$PC1), ylim=range(returns_PC_df$PC2))
```

There does not appear to be any clear separations between our stocks when projected into two dimensions.

#Clustering

Try hierarchical clustering with several different clustering methods.

First we compute a distance matrix.
```{r distance}
delta <- dist(returns %>% select(where(is.numeric)), method='DTW')
```

Try the Complete Linkage
```{r complete_linkage}
clust_complete <- hclust(delta, method='complete')
plot(clust_complete)
rect.hclust(clust_complete, k=5)
```

Try The Single Linkage
```{r single_linkage}
clust_single <- hclust(delta, method='single')
plot(clust_single)
rect.hclust(clust_single, k=5)
```

Try the Average Linkage
```{r average_linkage}
clust_average <- hclust(delta, method='average')
plot(clust_average)
rect.hclust(clust_average, k=5)
```
Try the Centroid Linkage
```{r single}
clust_centroid <- hclust(delta, method='centroid')
plot(clust_centroid)
rect.hclust(clust_centroid, k=5)
```

Try with 2 clusters

```{r clust}
clusters <- cutree(clust_complete, k=8)
ggplot(data=returns_PC_df, aes(x=PC1, y=PC2, color=as.factor(clusters))) + 
  geom_point() +
  scale_color_brewer(palette = "Accent") +
  coord_cartesian(xlim=range(returns_PC_df$PC1), ylim=range(returns_PC_df$PC2))
```

#Kmeans

```{r clust}
clusters <- kmedoids(delta, 3)
ggplot(data=returns_PC_df, aes(x=PC1, y=PC2, color=as.factor(clusters))) + 
  geom_point() +
  scale_color_brewer(palette = "Accent") +
  coord_cartesian(xlim=range(returns_PC_df$PC1), ylim=range(returns_PC_df$PC2))
```
